Spin qudit tomography and state reconstruction error

### 2021-12-16 sun_phases/qudit_tomo/qudit_tomo.tex
moving qudit tomo materials

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/qudit_tomo/}} % set path for all figures

@@

\begin{abstract}
We consider the task of performing quantum state tomography on a $d$-level spin qudit, using only measurements of spin projection onto different quantization axes.
After introducing a basis of operators closely related to the spherical harmonics, which obey the rotational symmetries of spin qudits, we map our quantum tomography task onto the classical problem of signal recovery on the sphere.
We then provide algorithms with $O\p{rd^3}$ serial runtime, parallelizable down to $O\p{rd^2}$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
Our algorithms motivate a simple randomized tomography protocol, for which we find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
\end{abstract}

@@

Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, lu2011strongly, mischuck2012control, aikawa2012boseeinstein, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, takekoshi2014ultracold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{hemmer2018squeezing, evrard2019enhanced}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence, perlin2021engineering}, and offer unique capabilities for quantum computation and error correction \cite{albert2020robust, gross2021designing, barnes2021assembly}.
In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.

@@

The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
Specifically, for any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can therefore define the rotation operator
\begin{align}
  R\p{\bm\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand rotated polarization operators as
\begin{align}
  T_{\bm\omega\ell m} \equiv
  R\p{\bm\omega} T_{\ell m} R\p{\bm\omega}^\dag
  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\bm\omega}}^* T_{\ell n},
  \label{eq:trans_rot}
\end{align}
where $\bar{\bm\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\bm\omega$, and
\begin{align}
  D_{mn}^\ell\p{\bar{\bm\omega}}
  \equiv \bk{\ell m|R\p{\bar{\bm\omega}}|\ell n}
  \label{eq:wigner_D}
\end{align}
are (Wigner) rotation matrix elements.
For reasons that will become clear shortly, throughout this work we will primarily consider rotations of the sphere that take the north pole to a point $\bm v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
For ease of notation, we therefore define $R\p{\bm v} \equiv R\p{\alpha,\beta,0}$, $T_{\bm v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\bm v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.

@@

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\bm v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere (see Figure \ref{fig:sphere_points}) \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\bm v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\bm v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\bm v}$ and associated coefficients $\tilde C_{\ell m}\p{\bm v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\bm v}$ at all points $\bm v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} f\p{\bm v}
  = \sum_{\bm v,k,n} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v}
  = \delta_{\ell k} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the the {\it reduced recovery problem}.
Due to the fact that $D^\ell_{0,m}\p{\bm v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\bm v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\bm v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\bm v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
A natural avenue to develop better spin qudit tomography protocols would therefore be to build on the existing classical signal recovery algorithms, tailoring them to solve the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} rather than the full recovery problem in Eq.~\eqref{eq:full_recovery}.
We leave these developments to future work.

@@

Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no grater than $\epsilon^2$:

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\bm v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
Save all measurement matrix data associated with the final measurement axes $V$ for later use.
\item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute estimates of $\bk{T_{\bm v\ell,0}}_\rho$,
\begin{align}
  \tilde T_{\bm v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu},
\end{align}
where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
\begin{align}
  Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  \p{x^V_{\ell k\bm v}}^* T_{\bm v\ell,0}, \\
  \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  x^V_{\ell k\bm v} \, \tilde T_{\bm v\ell,0},
\end{align}
and combine them into the estimate
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho.
\end{align}
\end{enumerate}
The expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}, which will additionally guarantee that $\tilde\rho_V$ satisfies all requirements for being a physical state.

@@

\begin{figure}
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Color indicates the qudit dimension $d$.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an arbitrarily chosen angle $\theta$.}
  \label{fig:axes}
\end{figure}

@@

For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$.
The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, namely the polar angle of all measurement axes, and provides no prescription for making this choice.
We therefore find the optimal choice of $\theta$ by minimizing the error scale $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1-\frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), which approaches $\pi/2$ as $d\to\infty$.
However, the error scale $\beta_{\pi/2}=\infty$, reflecting the fact that full state tomography is impossible with measurement axes lying in a single plane.
The method of Ref.~\cite{newton1968measurability} therefore requires extremely careful fine-tuning of measurement axis orientations for large spin dimensions.
For this reason, we expect our randomized tomography protocol be be more robust to errors in axis orientation.
We leave a detailed analysis of robustness to errors in axis orientation and the effect of these errors on state reconstruction to future work.

@@

Denoting the state of a spin-$s$ particle spin spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
\begin{align}
  S_\z \equiv \sum_{\mu=-s}^s \mu \op{s\mu},
  &&
  S_\pm \equiv \sum_{\mu=-s}^s
  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{s,\mu\pm1}{s\mu},
  \label{eq:spin_ops}
\end{align}
as well as
\begin{align}
  S_\x \equiv \f12\p{S_+ + S_-},
  &&
  S_\y \equiv -\f\i2\p{S_+-S_-},
  &&
  \bm S \equiv \p{S_\x,S_\y,S_\z}.
\end{align}
The spin vector $\bm S$ generates rotations of a spin-$s$ system in 3D space.
Specifically, the operator $e^{-\i\theta\bm S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.

>>>>

%%% figures
\usepackage{graphicx} % for figures
\graphicspath{{./figures/}} % set path for all figures

@@

\begin{abstract}
We consider the task of performing quantum state tomography on a $d$-level spin qudit, using only measurements of spin projection onto different quantization axes.
After introducing a basis of operators closely related to the spherical harmonics, which obey the rotational symmetries of spin qudits, we map our quantum tomography task onto the classical problem of signal recovery on the sphere.
We then provide algorithms with $O(rd^3)$ serial runtime, parallelizable down to $O(rd^2)$, for (i) computing a priori upper bounds on the expected error with which spin projection measurements along $r$ given axes can reconstruct an unknown qudit state, and (ii) estimating a posteriori the statistical error in a reconstructed state.
Our algorithms motivate a simple randomized tomography protocol, for which we find that using more measurement axes can yield substantial benefits that plateau after $r\approx3d$.
\end{abstract}

@@

Due to advancements in experimental capabilities to address nuclear spin states (i.e.~hyperfine levels) in ultracold atomic systems \cite{daley2011quantum, lu2011strongly, mischuck2012control, aikawa2012boseeinstein, smith2013quantum, cazalilla2014ultracold}, as well as developments in the control of ultracold molecular systems \cite{bohn2017cold, takekoshi2014ultracold, zeppenfeld2012sisyphus, kozyryev2017sisyphus, puri2017synthesis, wu2017cryofuge, marco2019degenerate, liu2019molecular, anderegg2019optical, chou2020frequencycomb, lin2020quantum}, a particular setting of growing interest is the spin qudit, or a multilevel quantum angular momentum degree of freedom.
Spin qudits can provide advantages over their qubit counterparts for quantum sensing \cite{evrard2019enhanced, hemmer2021squeezing}, enable quantum simulations of SU($d$) magnetism \cite{cazalilla2014ultracold, banerjee2013atomic, zhang2014spectroscopic, scazza2014observation, goban2018emergence, perlin2021engineering}, and offer unique capabilities for quantum computation and error correction \cite{albert2020robust, gross2021designing, barnes2021assembly}.
In all cases, quantum state tomography is necessary to take full advantage of a spin qudit\footnote{Note that the measurement of collective observables for quantum sensing or simulation can be recast as a single-spin tomography task.}.

@@

The polarization operators are spherical tensor operators, whose degree is preserved under 3D rotations generated by the spin operators $S_\x,S_\y,S_\z$.
Moreover, the degree-$\ell$ polarization operators $T_{\ell m}$ transform similarly to spin-$\ell$ particles and spherical harmonics $Y_{\ell m}$ under 3D rotations (see Appendix \ref{sec:rotations}).
Specifically, for any triplet of angles $\bm\omega=\p{\alpha,\beta,\gamma}$, we can define the rotation operator
\begin{align}
  R\p{\bm\omega} \equiv e^{-\i\alpha S_\z} e^{-\i\beta S_\y} e^{-\i\gamma S_\z},
\end{align}
and expand rotated polarization operators as
\begin{align}
  T_{\bm\omega\ell m} \equiv
  R\p{\bm\omega} T_{\ell m} R\p{\bm\omega}^\dag
  = \sum_{n=-\ell}^\ell D_{mn}^\ell\p{\bar{\bm\omega}}^* T_{\ell n},
  \label{eq:trans_rot}
\end{align}
where $\bar{\bm\omega}=\p{\gamma,\beta,\alpha}$ is the reversal of $\bm\omega$, and
\begin{align}
  D_{mn}^\ell\p{\bar{\bm\omega}}
  \equiv \bk{\ell m|R\p{\bar{\bm\omega}}|\ell n}
  \label{eq:wigner_D}
\end{align}
are (Wigner) rotation matrix elements.
For reasons that will become clear shortly, throughout this work we will primarily consider rotations of the sphere that take the north pole to a point $\bm v=\p{\alpha,\beta}$ at azimuthal angle $\alpha$ and polar angle $\beta$.
For ease of notation, we therefore define $R\p{\bm v} \equiv R\p{\alpha,\beta,0}$, $T_{\bm v\ell m} \equiv T_{\p{\alpha,\beta,0},\ell m}$, and $D^\ell_{mn}\p{\bm v} \equiv D^\ell_{mn}\p{0,\beta,\alpha}$.

@@

In fact, the problem of finding suitable axes $V$ and coefficients $C_{\ell m}\p{\bm v}$ to satisfy Eq.~\eqref{eq:tomo_recovery} can be mapped onto the well-studied problem of signal recovery on the sphere (see Figure \ref{fig:sphere_points}) \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality}.
The signal recovery problem can be stated as follows: given a square-integrable function $f$ on the sphere, with the spherical harmonic expansion
\begin{align}
  f\p{\bm v} = \sum_{\ell,m} f_{\ell m} Y_{\ell m}\p{\bm v},
\end{align}
where $f_{\ell m}$ are complex coefficients, find a set of points $V=\set{\bm v}$ and associated coefficients $\tilde C_{\ell m}\p{\bm v}$ with which we can reconstruct $f$, or equivalently its coefficients $f_{\ell m}$, from knowledge of the function's value $f\p{\bm v}$ at all points $\bm v\in V$; that is
\begin{align}
  f_{\ell m} = \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} f\p{\bm v}
  = \sum_{\bm v,k,n} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v} f_{kn}.
\end{align}
Reconstruction of functions with arbitrary coefficients $f_{\ell m}$ implies that
\begin{align}
  \sum_{\bm v} \tilde C_{\ell m}\p{\bm v} Y_{kn}\p{\bm v}
  = \delta_{\ell k} \delta_{mn},
  \label{eq:full_recovery}
\end{align}
which is a stronger version of the condition that we found for the spin qudit tomography problem in Eq.~\eqref{eq:tomo_recovery}.
We will refer to Eq.~\eqref{eq:full_recovery} as the {\it full recovery problem}, and Eq.~\eqref{eq:tomo_recovery} as the {\it reduced recovery problem}.
Due to the fact that $D^\ell_{0,m}\p{\bm v} = \sqrt{\frac{4\pi}{2\ell+1}}\, Y_{\ell m}\p{\bm v}$, any solution to the full recovery problem automatically solves the reduced recovery problem by setting $C_{\ell m}\p{\bm v} = \sqrt{\frac{2\ell+1}{4\pi}}\, \tilde C_{\ell m}\p{\bm v}^*$.
In principle, this mapping allows us to import a host of existing signal recovery algorithms \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} for the task of spin qudit tomography.
In practice, spin qudits typically have only a modest dimension $d$, which allows for simpler and optimized tomography protocols that are practical despite worse scaling with $d$ (see Section \ref{sec:protocol}).
A natural avenue to develop better spin qudit tomography protocols would therefore be to build on the existing classical signal recovery algorithms, tailoring them to solve the reduced recovery problem in Eq.~\eqref{eq:tomo_recovery} rather than the full recovery problem in Eq.~\eqref{eq:full_recovery}.
We leave these developments to future work.

@@

Assume, for example, that the estimates $\tilde T_j$ are equal to $\bk{T_j}_\rho$ up to uncorrelated noise with variance no greater than $\epsilon^2$:

@@

The ability to certify a statistical error bound on the empirical estimate $\tilde\rho_V$ of an unknown quantum state $\rho$ motivates the following protocol for spin qudit tomography:
\begin{enumerate}
\item Select a random set of measurement axes $V$ by uniformly sampling points on the sphere\footnote{To sample a point $\p{\alpha,\beta}$ from the uniform distribution on the sphere (with azimuthal angle $\alpha$ and polar angle $\beta$), you can sample a point $\p{a,b}\in[0,1]\times[0,1]$ from the uniform distribution on the unit square, and then set $\alpha=2\pi a$ and $\beta=\arccos\p{1-2b}$.}, and use any standard minimization algorithm to optimize the $2\abs{V}$ parameters in $V$ (two angles for each point $\bm v\in V$) by minimizing the quantum error scale $\epsilon_V$ in Eq.~\eqref{eq:bound}.
If $\abs{V}$ is too large for such optimization, you can simply generate many sets of random measurement axes, and then choose the set with the smallest quantum error scale $\epsilon_V$.
Note that computing the error scale $\epsilon_V$ requires, for each $\ell\in\set{0,1,\cdots,d-1}$, constructing the measurement matrix $M_{V\ell}$ in Eq.~\eqref{eq:block} and computing its singular value decomposition.
Save all measurement matrix data associated with the final measurement axes $V$ for later use.
\item For each axis $\bm v\in V$, make $n$ measurements of spin projection, and set $\tilde\Pi_{\bm v\mu}\approx\bk{\Pi_{\bm v\mu}}_\rho$ to the fraction of times in which the measurement outcome was $\mu$.
\item Use the estimates $\tilde\Pi_{\bm v\mu}$ of $\bk{\Pi_{\bm v\mu}}_\rho$ to compute estimates of $\bk{T_{\bm v\ell,0}}_\rho$,
\begin{align}
  \tilde T_{\bm v\ell,0}
  \equiv \sum_\mu \bk{\mu|T_{\ell,0}|\mu} \tilde\Pi_{\bm v\mu},
\end{align}
where the matrix elements of $T_{\ell,0}$ are provided in Eq.~\eqref{eq:trans_op}.
\item Denoting the nonzero singular values of $M_{V\ell}$ by $M^V_{\ell k}$ and the corresponding left singular vectors by $\bm x^V_{\ell k} = \sum_{\bm v} x^V_{\ell k \bm v} \ket{\bm v}$, compute the  operators and coefficients
\begin{align}
  Q_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  \p{x^V_{\ell k\bm v}}^* T_{\bm v\ell,0}, \\
  \tilde\rho^V_{\ell k} &\equiv \f1{M^V_{\ell k}} \sum_{\bm v}
  x^V_{\ell k\bm v} \, \tilde T_{\bm v\ell,0},
\end{align}
and combine them into the estimate
\begin{align}
  \tilde\rho_V = \sum_{\ell,k} \tilde\rho^V_{\ell k} Q_{\ell k}
  \approx \rho.
\end{align}
\end{enumerate}
The expected reconstruction error in $\tilde\rho_V$, or its root-mean-square distance from $\rho$, is provided by Eq.~\eqref{eq:error_cls}.
If measurement error is shot-noise-limited, then the error in $\tilde\rho_V$ is approximately $\E_V\p{\rho}\approx\E_V\p{\tilde\rho_V}$ and can be computed from Eq.~\eqref{eq:error_qnt}.
If $\tilde\rho_V$ has negative eigenvalues, its distance from $\rho$ can be reduced with maximum-likelihood corrections \cite{smolin2012efficient}, which will additionally guarantee that $\tilde\rho_V$ satisfies all requirements for being a physical state.

@@

\begin{figure}
  \centering
  \includegraphics{qudit_axes.pdf}
  \caption{Empirical measurement-adjusted error scales $\tilde\beta(p)$ with $p$ excess measurement axes, determined by minimizing over $10^3$ choices of measurement axes $V$ or 5 minutes of runtime (for each $p$), whichever comes first.
    Marker and color indicates the qudit dimension $d$.
    The rapid initial drop in $\tilde\beta(p)$ implies that using more measurement axes can substantially lower the upper bound on reconstruction error provided in Eq.~\eqref{eq:bound}, and that these benefits plateau after $p\approx d$.
    Horizontal reference lines mark the smallest measurement-adjusted error scales $\min_\theta\beta_\theta/\tilde\beta(0)$ achievable with the method in Ref.~\cite{newton1968measurability}, which is parameterized by an arbitrarily chosen angle $\theta$.}
  \label{fig:axes}
\end{figure}

@@

For reference, Figure \ref{fig:axes} also shows the smallest measurement-adjusted error scales $\beta_\theta$ achievable with the method of Ref.~\cite{newton1968measurability}, which is comparable to those achieved with our randomized protocol at $\abs{V}\approx3d$.
The method of Ref.~\cite{newton1968measurability} requires choosing an angle $\theta$, namely the polar angle of all measurement axes, and provides no prescription for making this choice.
We therefore find the optimal choice of $\theta$ by minimizing the error scale $\beta_\theta$ over all $\theta$ (see Appendix \ref{sec:angle_search}), and show $\min\beta_\theta/\tilde\beta(0)$ in Figure \ref{fig:axes}.
Empirically, we find that the optimal angle for the method of Ref.~\cite{newton1968measurability} is $\theta_{\t{opt}}\approx\frac{\pi}{2}(1-\frac1{1.34d})$ (see Appendix \ref{sec:angle_search}), which approaches $\pi/2$ as $d\to\infty$.
However, the error scale $\beta_{\pi/2}=\infty$, reflecting the fact that full state tomography is impossible with measurement axes lying in a single plane.
The method of Ref.~\cite{newton1968measurability} therefore requires extremely careful fine-tuning of measurement axis orientations for large spin dimensions.
For this reason, we expect our randomized tomography protocol to be more robust to errors in axis orientation.
We leave a detailed analysis of robustness to errors in axis orientation and the effect of these errors on state reconstruction to future work.

@@

Denoting the state of a spin-$s$ particle with spin projection $\mu$ onto a quantization axis by $\ket{s\mu}$, we define
\begin{align}
  S_\z \equiv \sum_{\mu=-s}^s \mu \op{s\mu},
  &&
  S_\pm \equiv \sum_{\mu=-s}^s
  \sqrt{s\p{s+1}-\mu\p{\mu\pm1}} \op{s,\mu\pm1}{s\mu},
  \label{eq:spin_ops}
\end{align}
as well as
\begin{align}
  S_\x \equiv \f12\p{S_+ + S_-},
  &&
  S_\y \equiv -\f\i2\p{S_+-S_-},
  &&
  \bm S \equiv \p{S_\x,S_\y,S_\z}.
\end{align}
The spin vector $\bm S$ generates rotations of a spin-$s$ system in 3D space.
Specifically, the operator $e^{-\i\theta\bm S\cdot\uv n}$ rotates a spin-$s$ system by an angle $\theta$ about the unit vector $\uv n$.
